Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-04-08 paper.tex
significant revisions

% Created 2017-04-07 Fri 16:52
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgad257ec}

@@

First proposed by \cite{Thorpe2005}, ``passthoughts'' authentication allow users to 
to submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) 
in a single step, by thinking a single mental task.
Even better, passthoughts have no externally visible ``tell,'' making them impervious to shoulder surfing attacks.
Passthoughts has been validated in a controlled environment using consumer-grade EEG devices with a single electrode \cite{Chuang2013b}, 
and using earpieces that collect EEG from the ear \cite{curranpassthoughts}.
\cite{Johnson2014} showed this protocol to be robust against impersonation attacks, even when the attack has learned the target's secret thought.
While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.

First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughst to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).

Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand how passthoughts change over time.
It will also help us understand the space of possible passthoguhts (Section 4).

@@

The main goal of this paper is to motivate and contextualize immediate studies that can address the above issues (Section 5).
The first study focuses on the use of passthoughts in a realistic, though lab-constrained setting, 
allowing a first look at how passthoughts might work in a real-time application context.
The second proposed study focuses on people's longitudinal relationships with EEG in their everyday life,
focusing not on behavioral differences experimental conditions, but on evolving attitudes (and signals) over longer timescales,
allowing assessment of both dynamic user attitudes, and the possible impact of shifting neural signals.

Passthought authentication stands a chance at becoming the first brain-computer interface to reach wider adoption. 
As such, passthoughts promises not only more usable, multi-factor authentication,
but also a source of data for those attempting to build other brain-computer interfaces,
The strategy is not without risks (Section 6),
but, if future work can address the two challenges posed here, several interesting paths unfurl.
e.g. for people with motor disabilities \cite{Mattia2013}.
closed-loop authentication, continuous and/or tacit authentication, 
and possible theoretical contributions to neuroscience, authentication and HCI (Section 7).

\section{Background}
\label{sec:org4dc5ef5}

@@

An ongoing problem in authentication lies in balancing strong security
(i.e., combining multiple factors of authentication)
with usability.
As an example, major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge factor),
and subsequently receives a code on their cellphone (a posession factor).
However, submitting two different 
authenticators in two separate steps has frustrated wide adoption
due to its additional hassle to users. The Apple iPhone, for instance,
already supports device unlock using either a user-selected passcode or a fingerprint. The
device could very well support a two-step two-factor authentication scheme if
desired. However, it is easy to understand why users would balk at having to
enter a passcode \emph{and} provide a fingerprint each time they want to unlock their phone.

\subsection{One-step, multi-factor authentication}
\label{sec:org416ad55}

@@

Authenticators are susceptible to a variety of attacks. 
One common attack on passwords is ``shoulder surfing,'' in which an attack uses visual or other cuesto steal, or improve the chances of guessing, a target's chosen secret. 
Passthoughts mitigates this attack by nature of the mental gesture:
since the expression of a passthought is not externally visible, the authenticator is impervious to shoulder surfing attacks.

\subsection{Passthought authentication}
\label{sec:org9645597}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgfd91d74}

@@

\subsection{Contending with mind-reading machines}
\label{sec:orgdc80023}

Finally, a crucial thread of past work work concerns how users perceive the capabilites of devices that purport to ``read'' their ``mind.''
Biosensing devices in general raise many questions for consumers.
You might be eligible for an insurance discount if you wear a FitBit \cite{Bernard2015} (depending, of course, on what readings the FitBit produces \cite{Brain2015}). 
But, would you wear a device in the workplace \cite{solon2015}, if your manager used it to track your productivity?
If biosensor data can be used in the courtroom \cite{Crawford2014}, could not pervasive biosensing help to \emph{predict} crime \cite{Thompson2011}? 
After all, one study suggests that probability of involvement in violent crime can be predicted from one's resting heartrate \cite{Latvala2015}.

In all of these examples, biosensing technologies blur the line between \emph{sensing bodies} and \emph{sensing minds}. 
Now, when people decide to buy sensor-equipped consumer devices \cite{Stables2016}, or get sensed passively by devices integrated into the walls and ceilings \cite{Adib2015} or city streets \cite{Thrift2014}, end-users will need to contend with the prospect of mind-reading machines.

@@

Crucially, there are some people who actually \emph{want} their minds measured, e.g. for self-reflection. Consider the Spire, a breath sensor that claims to divine, from a person's patterns of in-breaths and out-breaths, what the user is calm, focused, or tense.
For the device to ``work,'' not only must these detected signals match with end-users' intuitions, but users must also believe that a device like the Spire has the power to measure and detect these phenomena, given breath as input \cite{Ali2014a}. 
In general, technologies that claim to ``measure the mind'' must rely on end-users to define the criteria by which systems are deemed effective, or accurate.

If we wish to understand what role passthought authentication \emph{could} play in day-to-day life,
we must view it both through the lens of potential privacy concerns, \emph{and} through the lens of possible opportunities for self-reflection and self-understanding. 
Of course, users' attitudes will not be fixed: they will evolve over time, as users observe the device in action, and correlate its judgments with their own lived experiences \cite{Nafus2016}.

\section{User Perceptions of EEG}
\label{sec:org30e9f5a}

The prior section outlined the first major challenge to passthought authentication: that of corpus diversity.
The following section reviews a more subtle challenge: that of usability, as it relates to attitudes around sensing brainwaves.

What can machines know about a person's mind, even theoretically? 
This question is never more relevant than when speaking of attempts at quantitative measures of brain activity (e.g. brainwaves).
Past work has established the almost magical abilities that people tend to ascribe to brain-scanning devices, even subjects with specific training in the limitations of brain-scanners \cite{Ali2014a}.
This section outlines concerns around ``mind-reading'' machines, and how they relate to EEG and passthoughts specifically.
We then move to a discussion for possible ways to address these concerns, and concerns about dataset diversity, in the following section.

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org4244390}

In our preliminary findings, brainwaves (EEG) are seen as among the most revealing biosignals, just below body language and facial expression, in their capacity to reveal the goings on of a person's mind. 
More common sensors such as GPS and step count are seen as less revealing (despite empirical evidence suggesting such data can be quite revealing indeed \cite{Canzian2015}).

The survey I report on here, currently in-progress, examines how people's beliefs differ given device ownership, and their membership in one of two groups: Mechanical Turk workers, or people enrolled in Health-e-Heart, a massive (n > 40,000), longitudinal study, in which volunteers fill out surveys about themselves, and/or upload data from biomedical self-tracking devices, over the course of several years \cite{Estrin2010a}.
In one portion of the survey, we ask subjects to rate a number of different biosensors in order of how likely individual's believe each sensor is to reveal what ``a person is thinking or feeling'' (Figure \ref{fig:rank}).

@@

What will this finding mean for wider adoption? 
Will people shy away from using their passthought authenticator in certain situations, or when they are feeling some type of way?
The following section describes one exploratory study that could investigate some aspects of this larger question.

\section{Diversity and crackability of passthoughts}
\label{sec:orgba47d6a}

To transition from the lab to the real world,
passthoughts studies must collect larger, and more diverse corpora of EEG data.
While past work on passthoughts has achieved excellent results on corpora of recordings from different users, 
these studies do not consider passthoughts from a variety of different subject conditions.
For example, sitting subjects may have different patterns of neural activity from subjects who are standing, walking or exercising \cite{Thibault2016a},
let alone subjects who are under the influence of caffeine, alcohol or marijuana.

Relatedly, these studies do not systematically investigate how these recordings relate statistically to non-passthoughts.
That is, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If I have a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.
Investigating this question could also help us understand how and why passthoughts work at all: Why are passthoughts unique, and how unique are they?

\section{Two studies on passthoughts}
\label{sec:org31b107a}

This section proposes two studies on passthought authentication.
One study, a controlled, lab-based experiment, seeks to raise fundamental questions about how the feedback fo a real-time authentication system may affect the performance of passthoughts.
It also begins to address certain, limited questions around the shifting nature of neural signals.
The second study, a longitudinal deployment, seeks to collect a large and diverse corpus of EEG signals, while probing people's beliefs and attitudes about EEG and brainscanning in everyday life.
Together, the findings of these proposed studies could serve as a starting point for addressing the major questions raised in the prior sections.

\subsection{A real-time passthought authenticator}
\label{sec:orgf895ad6}

Past work has proposed passthoughts as a more usable form of multi-factor authentication,
as compared to existing protocols.
However, no study yet has systematically evaluated this claim.
Here, we propose a study aimed at 
examining passthought's usability in an ecologically valid context.

This study would take place in a lab, under the supervision of an experimenter.
First, the experimenter would calibrate a subject with a passthought authenticator, as in \cite{Chuang2013b}.
Through an automated cross-validation process, the participant's best-performing passthought would be selected.

Next, the experimenter sould present users with an online banking application, and ask them to perform their passthoughts.
We can manipulate feedback such that users either see the real authentication accuracy (control), 
are always rejected by the authenticator, 
or always accepted by the authenticator.

\subsubsection{The effect of feedback}
\label{sec:orgc529a54}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org6714417}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgc659ab1}

The study proposed above would help answer preliminary questions about
the usability of a passthought authenticator,
and possible ways for dealing with shifting neural signals.
However, as reviewed above, two major challenges would still remain for passthoughts authentication.

First, we do not have a sufficiently large corpus of EEG signals, 
preventing us from investigating how robust passthoughts authentication performs in various user conditions,
and from understanding how easy particular passthoughts are to guess or crack.
Second, we do not understand how people's beliefs about EEG might affect their behavior with a passthought authentication system.
Unfortunately, these challenges make it difficult to produce a passthought authenticator that works with any reliability in real contexts,
making, a longitudinal study with a working authenticator not immediately practical.

However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
A technology probe or diary study \cite{Gaver1999} could help address both of these issues at the same time

A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

The following two sections expand on both aspects of this project in greater detail.

\subsubsection{A more diverse corpus}
\label{sec:orgb90c729}

@@

In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as tSNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org4421f28}

@@

Of course, this study is no substitute for a working, online passthoughts authentication system.
Instead, this study aims to collect useful data before such a system exists.
It will not only elicit beliefs, 
but also allow us to collect larger datasets, 
and to catch technical issues in sensing devices and collection platforms.

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org05fd2a6}

After the study described above, future work should be able to start building more robust, world-ready passthought systems,
which could offer improvements to the usability and security of authenticating.
However, these opportunities do not come without risks. 
Indeed, some risks are unique to the application context, and to EEG as a class of biosignal. 
This section briefly reviews risks to user privacy and security that widespread passthought authentication may introduce. 
I do not pose specific challenges to passthoughts here, though many surely lurk; instead, I present broad class of categories from which questions may emerge.

\subsection{Privacy}
\label{sec:org665d86b}
One clear risk comes to user privacy, as it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could, at least theoretically, come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular datbase queries.

\subsection{Security}
\label{sec:orgf04f8ff}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

There is also the question of the security of large data repositories in which EEG data might be stored.
Large data repositories are what Wolf \cite{Wolf2010} calls a ``toxic asset''; infrastructures that must be maintained, lest the maintainer take liability for the potentially harmful fallout of poor data management.
With biosignals, as with many kinds of data, it is not entirely clear what they might mean until they are already collected in aggregate. 
At this point, it is too late to decide on an appropriate data security policy.
Good data encryption policies should be built into collection systems from the very beginning,

It remains an open question what specific protections and access controls will yield robust security.
Homomorphic encryption, in which computation such as database queries can be performed on encrypted data, provides one interesting path for future work \cite{Tu2013}.

\section{Further Future Directions}
\label{sec:org8adb539}

This paper so far has motivated a longitudinal study with EEG, and its importance even before a working passthought authenticator has been completed. 
I have also discussed potential risks intrinsic to the development of passthoughts systems.
With these risks in mind, the present section explores some of the exciting possibilities that could open up after the immediate priorities described previously.

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:orgedc89f4}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to pre-empt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:orgd885dff}

>>>>

% Created 2017-04-08 Sat 12:10
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org2972ffd}

@@

First proposed by \cite{Thorpe2005}, ``passthoughts'' authentication allow users to 
to submit both a knowledge factor (i.e., secret thought) and an inherence factor (i.e., brainwave signal unique to the individual) 
in a single step, by performing a single mental task.
Since this original proposal, passthoughts has been validated in a controlled environment using consumer-grade devices \cite{Chuang2013b}, 
and using in-ear EEG earbuds \cite{curranpassthoughts}.
\cite{Johnson2014} showed this protocol to be robust against impersonation attacks, even when the attack has learned the target's secret thought.

While recent work is heartening, passthoughts remains confined, for now, to the lab.
This paper reviews the immediate questions that must be addressed if passthoughts is to move into everyday life.

First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughts to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).

Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand how passthoughts change over time.
It will also help us understand the space of possible passthoughts (Section 4).

@@

The main goal of this paper is to motivate and contextualize studies that can begin to address these challenges (Section 5).
The first study focuses on the use of passthoughts in a realistic, though lab-constrained setting, 
allowing a first look at how passthoughts might work in a real-time application context.
The second proposed study focuses on people's longitudinal relationships with EEG in their everyday life,
focusing not on behavioral differences experimental conditions, but on evolving attitudes (and signals) over longer timescales,
allowing assessment of both dynamic user attitudes, and the possible impact of shifting neural signals.

Passthought authentication stands a chance at becoming the first brain-computer interface to reach wider adoption. 
As such, passthoughts promises not only more usable multi-factor authentication,
but also a source of data for future development of EEG-based brain-computer interfaces
e.g. for people with motor disabilities \cite{Mattia2013}.
Passthoughts is not without risks (Section 6),
but, if future work can address the two challenges posed here, several interesting paths unfurl.
closed-loop authentication, continuous and/or tacit authentication, 
and possible theoretical contributions to neuroscience, authentication and HCI (Section 7).

\section{Background}
\label{sec:org841d0eb}

@@

An ongoing problem in authentication lies in balancing strong security
(i.e., combining multiple factors of authentication)
with usability.
As an example, major industry players such as Google and
Facebook have strongly encouraged their users to adopt two-factor
authentication (2FA), in which a user enters his or her password (a knowledge factor),
and subsequently receives a code on their cellphone (a posession factor).

However, submitting two different 
authenticators in two separate steps has frustrated wide adoption
due to its additional hassle to users. The Apple iPhone, for instance,
already supports device unlock using either a user-selected passcode or a fingerprint. The
device could very well support a two-step two-factor authentication scheme if
desired. However, it is easy to understand why users would balk at having to
enter a passcode \emph{and} provide a fingerprint each time they want to unlock their phone.

\subsection{One-step, multi-factor authentication}
\label{sec:org578b4c9}

@@

Authenticators are susceptible to a variety of attacks. 
One common attack on passwords is ``shoulder surfing,'' in which an attack uses visual or other cues to steal, or improve the chances of guessing, a target's chosen secret. 
Passthoughts mitigates this attack by nature of the mental gesture:
since the expression of a passthought is not externally visible, the authenticator is impervious to shoulder surfing attacks.

\subsection{Passthought authentication}
\label{sec:org6be6687}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgae7c8cc}

@@

\section{User attitudes and perceptions}
\label{sec:orgcd9b457}

In spite of past work making passthoughts less visible with more convenient form-factors,
a large question still remains:
What sense would people make of passthoughts, as a technology, in their everyday life?
This question begs not only user-centered design studies with passthoughts itself,
but more general questions about what EEG means to people,
and what people believe EEG data can reveal about them.
Past work has established the almost magical abilities that people tend to ascribe to brain-scanning devices, even subjects with specific training in the limitations of brain-scanners \cite{Ali2014a}.
This section outlines common concerns around ``mind-reading'' machines, and how they relate to EEG and passthoughts specifically.

\subsection{Contending with mind-reading machines}
\label{sec:org6713e76}

Biosensing devices in general raise many questions about privacy for end-users,
typically around the meaning of the data produced by particular devices.
For example, you might be eligible for an insurance discount if you wear a FitBit \cite{Bernard2015} (depending, of course, on what readings the FitBit produces \cite{Brain2015}). 
But, would you wear a device in the workplace \cite{solon2015}, if your manager used it to track your productivity?
If biosensor data can be used in the courtroom \cite{Crawford2014}, could not pervasive biosensing help to \emph{predict} crime \cite{Thompson2011}? 
After all, one study suggests that probability of involvement in violent crime can be predicted from one's resting heartrate \cite{Latvala2015}. 
In all of these examples, biosensing technologies blur the line between \emph{sensing bodies} and \emph{sensing minds}. 
Now, when people decide to buy sensor-equipped consumer devices \cite{Stables2016}, or get sensed passively by devices integrated into the walls and ceilings \cite{Adib2015} or city streets \cite{Thrift2014}, end-users will need to contend with the prospect of mind-reading machines.

@@

If we wish to understand what role passthought authentication \emph{could} play in day-to-day life,
we must view it both through the lens of potential privacy concerns, \emph{and} through the lens of possible opportunities for self-reflection and self-understanding. 
Of course, users' attitudes will not be fixed: they will evolve over time, as users observe the device in action, and correlate its judgments with their own lived experiences \cite{Nafus2016}.
In the next section, we turn a discussion on EEG specifically as a way of motivating questions around the meaning people may build around passthought authenticators.

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org8e44d66}

The survey we report on here, currently in-progress, examines how people's beliefs differ given device ownership, and their membership in one of two groups: Mechanical Turk workers, or people enrolled in Health-e-Heart, a massive (n > 40,000), longitudinal study, in which volunteers fill out surveys about themselves, and/or upload data from biomedical self-tracking devices, over the course of several years \cite{Estrin2010a}.
In one portion of the survey, we ask subjects to rate a number of different biosensors in order of how likely individual's believe each sensor is to reveal what ``a person is thinking or feeling'' (Figure \ref{fig:rank}).

In our preliminary findings, brainwaves (EEG) are seen as among the most revealing biosignals, just below body language and facial expression, in their capacity to reveal the goings on of a person's mind. 
More common sensors such as GPS and step count are seen as less revealing (despite empirical evidence suggesting such data can be quite revealing indeed \cite{Canzian2015}).
What will this finding mean for wider adoption? 
Will people shy away from using their passthought authenticator in certain situations, or when they are feeling some type of way?

@@

\section{Diversity and security of passthoughts}
\label{sec:orge34b851}

While the previous section outlined questions around user attitudes, empirical questions about passthoughts, as signals, also linger.
This section outlines and motivates the major quantitative questions that have not been fully answered by past work on passthoughts.

While past work on passthoughts has achieved excellent results on corpora of recordings from different users, 
these studies do not consider passthoughts from a variety of different subject conditions.
Passthoughts studies must collect larger, and more diverse corpora of EEG data to examine how EEG signals change (or remain stable) throughout the dynamic contexts of daily life.
For example, sitting subjects may have different patterns of neural activity from subjects who are standing, walking or exercising \cite{Thibault2016a},
let alone subjects who are under the influence of e.g. caffiene or alcohol.

Investigating this question could also help us understand how and why passthoughts work at all: Why are passthoughts unique, and how unique are they?
A primary question in passthoughts surrounds how large the real space of possible passthoughts might be \cite{Thorpe2005}.
While the space of possible passthoughts is potentially unlimited, we do not yet know what passthoughts we stand a reasonable chance at observing consistently over time.
A larger corpus of data might help shed light on this issue by allowing us to observe the distribution of signals that people produce over time.

A more subtle, but closely related question, surrounds how passthought EEG recordings relate statistically to non-passthought EEG recordings.
In other words, we do not know how the particular passthoughts observed in past work are drawn from the distribution of EEG signals that an individual produces over the course of their day.
This blind-spot poses a possible challenge to passthought's vulnerability to dictionary-style cracking.
If an attacker has a large enough corpus of EEG readings, do some passthoughts start to look as guessable as \emph{password1234}?
By answering such questions, we could design data-driven policies for, e.g., how many retry attempts passthought authenticators should allow.

\section{Two studies on passthoughts}
\label{sec:org649bbe2}

The prior two sections raise two main topics surrounding passthoughts that future work could address. 
First, our limited understanding of passthoughts' usability, and user attitudes about the sensing modality present immediate questions for further development of this technology.
Meanwhile, our limited knowledge of how passthoughts shift and change over time, and around the diversity of EEG signals as our limited (non-medical) devices sense them,
raise questions about how frequently passthoughts would need to be calibrated, how accurate we can expect the protocol to be in different context, and how secure it might remain under threat from a motivated attacker.

This section proposes two studies on passthought authentication which, taken together, could make headway on these topics.
One study, a controlled, lab-based experiment, seeks to raise fundamental questions about how the feedback of a real-time authentication system may affect the performance of passthoughts.
It also begins to address certain, limited questions around the shifting nature of neural signals.
The second study, a longitudinal deployment, seeks to collect a large and diverse corpus of EEG signals, while probing people's beliefs and attitudes about EEG and brainscanning in everyday life.
Together, these studies address both long-term concerns about user attitudes and signal diversities, and also short-timescale questions about the usability and accuracy of passthoughts in realistic use scenarios.

\subsection{A real-time passthought authenticator}
\label{sec:org852f286}

Passthoughts is supposed to be a more usable form of multi-factor authentication
as compared to existing protocols,
as it provides both a knowledge and an inherence factor in a single step.
However, no study yet has systematically evaluated passthoughts' usability.
Here, we propose a study aimed at examining passthoughts' usability in an ecologically valid context.

\subsubsection{Study protocol}
\label{sec:org44248b2}

This study would take place in a lab, under the supervision of an experimenter.
First, the experimenter would calibrate a subject with a passthought authenticator, as in \cite{Chuang2013b}.
Through an automated cross-validation process, the participant's best-performing passthought would be selected.
Next, the experimenter would present users with an online banking application, and ask them to perform their passthoughts.
We can manipulate feedback such that users either see the real authentication accuracy (control), 
are always rejected by the authenticator, 
or always accepted by the authenticator.

After this task, subjects could take a post-questionnaire including various usability questions.
After filling out this questionnaire, the experimenter might engage users in a brief, ten-minute semi-structured interview,
in which subjects were asked to recount their experience with the authenticator, in their own words.
This interview could help gain some richer, qualitative data that traditional survey methods might fail to capture.

\subsubsection{The effect of feedback}
\label{sec:org7cdb796}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgfd1298c}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgc109281}

The study proposed above would help answer preliminary questions about
the usability of a passthought authenticator in a short-term context,
and possible ways for dealing with shifting neural signals,
a few questions will still remain.
First, the study above will not help us collect a large corpus of EEG signals, 
preventing us from investigating how robust passthoughts authentication performs in various user conditions,
and from understanding how easy particular passthoughts are to guess or crack.
Second, while the previous study helps us understand user attitudes over a short timescale,
it will not help us understand how people's beliefs about EEG might change over longer periods of time, as they use their devices in day-to-day life.

Unfortunately, these challenges - particularly those around shifting neural signals - also make it difficult to produce a passthought authenticator that works with any reliability in real contexts.
This makes a longitudinal study with a working authenticator impractical for now.
However, we may still perform a longitudinal study that allows us to interrogate the usability aspects around (and attitudes about) passthoughts specifically, and EEG generally.
In so doing, we may also collect a larger and more diverse corpus of passthoughts, which can be used to address the paucity of data we face today.
A technology probe or diary study \cite{Gaver1999} could help address both of these issues at the same time.

Of course, this study would be no substitute for a working, online passthoughts authentication system.
Instead, this study aims to collect useful data before such a system exists.
It will not only elicit beliefs, 
but also allow us to collect larger datasets, 
and to catch technical issues in sensing devices and collection platforms.

\subsubsection{Study protocol}
\label{sec:org25d4283}
A small group of subjects could wear a working, recording EEG device, whether or not it provides feedback, in a variety of settings for some number of days,
having subjects journal their experiences and asking them specifically what they feel someone might be able to know about them from the EEG signals they record.
At the same time, we could use this study as an opportunity to collect a much larger, and more diverse corpus.
To aid in the collection of signals that are specific to our problem of passthought authentication,
subjects in this study might be prompted to perform a variety of tasks at a few pre-chosen points throughout the day.
With the data collected during this study, we could easily simulate passthought accuracy on a much more realistic (and representative) sample of readings.

@@

\subsubsection{A more diverse corpus}
\label{sec:orga8125e2}

@@

\subsubsection{The space of possible passthoughts}
\label{sec:org8672aa4}
In another potentially fruitful analysis, such a corpus will allow us to perform statistical analysis of how passthoughts are drawn from the overall distribution of EEG signals. 
Using multi-dimensional clustering algorithms such as t-SNE \cite{VanDerMaaten2008} 
could assist us in understanding how particular passthoughts relate to other EEG signals that an individual expresses involuntarily throughout the day. 
These clusters will help us understand how rare or unlikely a given passthought is, and help shed light on why and how given passthoughts are expressed uniquely between individuals.

@@

\subsubsection{Usability and attitudes}
\label{sec:org6bec26f}

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org210fd82}

After the studies described above, 
we will have a much better grasp on the usability, and security properties of passthought authentication.
However, there may still be unexplored risks, challenges, and tradeoffs,
especially around user privacy.
Indeed, some of these risks are unique to the application context, and to EEG as a class of biosignal. 
This section briefly reviews risks to user privacy and security that widespread passthought authentication may introduce. 
We present broad class of categories from which such risks may emerge.

\subsection{Privacy}
\label{sec:orgabf551f}
As of yet, it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular database queries.

\subsection{Security}
\label{sec:org5785718}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

There is also the question of the security of data infrastructures in which EEG data might be stored.
Large data repositories are what Wolf \cite{Wolf2010} calls a ``toxic asset''; they must be maintained, 
lest the maintainer take liability for the potentially harmful fallout of poor data management.
With biosignals, as with many kinds of data, it is not entirely clear what they might mean until they are already collected in aggregate. 
At this point, it is too late to decide on an appropriate data security policy.

Strong encryption policies should be built into collection systems from the very beginning, 
It remains an open question what specific protections and access controls will yield robust security.
Homomorphic encryption, in which computation such as database queries can be performed on encrypted data, provides one interesting path for future work \cite{Tu2013}.

\subsection{Tradeoffs between security and privacy}
\label{sec:orgc48b801}

In some cases, passthoughts could present direct tradeoffs between security and privacy.
For example, end-user privacy could be enhanced by storing all data locally, on the phone. 
All classification, and the training of all classifiers, could occur locally, so that users never need to disclose their private biosensory data to a third party.
However, security might be improved by aggregating user data so as to construct more robust, reliable classifiers.
Aside from classifier accuracy, training classifiers in the cloud could help with the speed of calibration,
and prevent undue battery drain on user devices.

These factors suggest a possible tension between the accuracy (and thus security) of passthought authentication,
and the locality (and thus privacy) of potentially sensitive user data.
Future work should explore this tradeoff empirically, using real data and simulations from a variety of different users.
Future work might also explore metrics by which to judge such tradeoffs.
Whereas security might be measured straightforwardly using false-acceptance and false-rejection rates,
user privacy might be more challenging to quantify, as might the tradeoffs between the two.
However, future work will need to address these issues if we are to balance users' security requirements with their privacy requirements.

\section{Further Future Directions}
\label{sec:org181ef3b}

This paper so far has motivated two future studies on passthoughts,
and discussed potential risks intrinsic to the development of passthoughts systems.
With these risks in mind, the present section explores some of the exciting possibilities that could unfold after the immediate priorities described in the prior sections.

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:org7ffd58a}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to preempt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:org9ee5756}
